BOLD Parameter Estimation Using Particle Filters

### 2010-07-22 doc/thesis/results.tex
fixed up results.

\chapter{Simulation}
\label{sec:SimulationResults}
Two levels of simulation are discussed in this section. First the simulation of single time series
and second, a slice of simulated FMRI time series. The single time series tests
are necessary to investigate the power of particle filters in 
identifying the BOLD model in a noisy environment. If the particle filter is unable
to get back to the \emph{true} time-series in spite of noise then the particle filter
will not be useful. Additionally, determination of reasonable parameters from the
time-series is also desirable. The reason to then simulate an entire slice is that
Physics-Oriented Simulated Scanner for Understanding MRI (POSSUM)
models noise more realistically, for instance by noise in the frequency domain. Additionally 
it provides insight into the applicability of particle filters in modeling the BOLD 
signal on a large scale. Both sorts of tests are usefull in determining the
consistency of the particle filter. For the single voxel tests, each test will
actually be performed eleven different times, each with a new noise realization.
If runs fail to have consistent estimates of the underlying BOLD signal, then
this would indicate a large variance error in the model. Similarly, in the
Possum simulation there are actually only four discrete parameter sets driving the time-series.
Nevertheless each voxel will have a different tissue composition and a different
noise realization. Therefore the Possum simulation should also provide a good
test of the consistency of the BOLD model.

\section{Single Time Series}
\label{sec:Single Voxel Simulation}
Given the state-space equations for the BOLD signal, simulating a single time 
series is relatively straight forward. After a true signal is generated,
identically and independently distributed (I.I.D.) Gaussian noise and a Wiener
process with Gaussian I.I.D. steps are added to the true signal. Finally a 
carrier level is added, since BOLD is typically
measured as a \% difference from the baseline. The particle filter
algorithm will immediately remove this by calculating the \% difference, 
but adding a carrier level means that the exact same algorithm which 
will is used for simulated data may be used for the real data.

Once this noisy simulated time series has been generated, the particle filter algorithm
 is then run on this single time series image. Here I will include two sets of tests 
to determine the power of the particle filter in modeling. The first set are 
used to demonstrate the ability of the particle filter to find the most likely
set of parameters/state variables over the course of the run. The second type
of test will investigate the problem of false-positives. As the first
round of tests will show, given that there is an underlying BOLD signal, 
the particle filter is excellent at finding the most probable cause of 
the observed signal. However, as discussed in the false positive tests,
even if there is no true underlying signal, the particle filter will still determine
the most likely BOLD signal; which may not be flat. Thus, it is 
important to investigate methods of determining false positives using 
average residuals. This problem could in actuality never come up, for instance
if the number of measurements is very large or if the noise peaks are not
particularly large.

%LOW NOISE SECTION, with a signal
\subsection{Signal with Low Noise}
\label{sec:SimLowNoise}
%tests had noise of $ \{\sigma_y = .01, \sigma_x = .005\} $, where $\sigma_y$ is the
%measurement noise, and $\sigma_x$ is the wiener step size. Both signals used the
%parameters .
%The particle filter used the parameters defined in \autoref{tab:Prior} (\autoref{sec:PriorDistrib}),
%thus the particle filter was not centered over the correct values. 
\begin{figure}
\includegraphics[clip=true,trim=6cm 2cm 6cm 3.5cm,width=17cm]{images/realization_lownoise}
\caption{Test Signals with low noise compared to the clean signal.}
\label{fig:LowNoiseRealization}
\end{figure}

\begin{figure}
\includegraphics[clip=true,trim=6cm 2cm 6cm 3.5cm,width=17cm]{images/preprocessed_lownoise}
\caption{A comparison of the preprocessed signals for the low noise case. This is the
noisy input to the actual particle filter algorithm.}
\label{fig:PreprocessedLowNoise}
\end{figure}

\begin{figure}
\includegraphics[clip=true,trim=6cm 2cm 6cm 3.5cm,width=17cm]{images/comparison_lownoise}
\caption{A comparison of the fitted signals for the low noise case.}
\label{fig:FitComparisonLowNoise}
\end{figure}

To begin the single voxel simulation; I generated a signal using the following parameters:
$\{\tau_0 = 1.45, \alpha = .3, E_0 = .47, V_0 = .044, \tau_s = 1.94, \tau_f = 1.99, \epsilon = 1.8\}$.
These same parameters will be used again in the high noise case. The noise realizations were
then based on a measurement noise, $\sigma_y$, of $.001$ and a drift standard deviation,
$\sigma_x$ of $.0005$. The measurement noise as well as the steps of the drift
were taken to be Gaussian
For this low noise case, the eleven realizations are shown in \autoref{fig:LowNoiseRealization}.
The actual signal that was delivered into the particle filter
is the result of some preprocessing to remove drift, as described in 
\autoref{sec:Methods Preprocessing}. The bias introduced into the signal by preprocessing 
may have some effect on the resulting fit; thus the preprocessed signal is compared
to the true BOLD signal in \autoref{fig:PreprocessedLowNoise}.
The preprocessing consists of several steps, which are discussed in detail in \autoref{sec:Methods Preprocessing}.
Obviously the spline struggles to match the signal near the end, as the graph shows, although 
overall, the preprocessing seems to have been successful at removing trends. Finally, the 
set of fits to the true BOLD signal are shown in \autoref{fig:FitComparisonLowNoise}.
The filtering property of the particle filter is clear from the results here.  In several locations the output
of the particle filter looks like a filtered version of the input. For instance toward the
end, the estimates stay flat in spite of the preprocessed data drifting off. By
this point, the algorithm has converged sufficiently to know that that such a movement in
signal isn't possible. A similar circumstance occurs at 100 seconds in. 
In general the fits are good, although the bias introduced by pre-processing 
results in a slightly larger peak signal. The final parameter sets are shown in 
\autoref{tab:LowNoiseResults}. Note that throughout the results, unless otherwise specified,
I will define the residual as the square root of the mean squared residual, 
\begin{equation}
\text{Residual} = \sqrt{\sum (\hat{y}_k - y_k)^2}
\end{equation}
where $\hat{y}_k$ is the estimated output at time $k$ and $y_k$ is the output
sample at time $k$. Note the subtle distinction between this and the error
which I will define as 
\begin{equation}
\text{Error} = \sqrt{\sum (\hat{y}_k - Y_k)^2}
\end{equation}
where $\hat{y}_k$ is the estimated output, and $Y_k$ is the underlying (free of noise)
output. Throughout this section and the next, error and residual of a time series 
will be referencing these values.

@@

There are a few results worth noting here. First the time constants vary greatly across
runs with different noise realizations, yet the sum of the individual time constants
($\tau_f$, $\tau_s$ and $\tau_0$) seems to be more consistent. In general the 
time constants are falling short of the true time constant. This could be a limitation
based on the prior distribution (which notably has an initial mean below the true values) 
or it could be caused by the disassociated benefits of a correct time
constant with correct output. It often takes several measurement periods before a difference
in time constants becomes relevant to the weight. Its also possible that output is insensitive
to changes in the time constants and that varying the activation durations would 
cause the variations to be more visible. Another interesting result in the huge variation 
in the levels of $V_0$. In general,
with this admittedly small amount of noise, it would appear that the relation between a set
of parameters/stimuli and the output is not injective. In other words,  a time
series is not unique to a single set of parameters. This is good justification that 
simultaneous blood volume or tagged flow calculations with the conventional FMRI 
could benefit the model.

The covariance matrix (\autoref{tab:CovSim}) also confirms the idea that the 
$\tau$ parameters are interchangeable
as far as the BOLD signal is concerned. Notice the covariance of $\tau_f$ and $\tau_0$
is $-0.019$ whereas
the variance of $\tau_0$ and $\tau_f$ are $0.019$ and $0.04$ respectively. Clearly there is 
some measure of ill-defined behavior between the different time-constants, as well as several
other variables.  The convergence properties of the first run in \autoref{tab:LowNoiseResults} 
demonstrates the migration of parameters to match the estimates. Given the relatively high
variance of parameters, its likely that more measurements or perhaps more variation
in the stimuli could further differentiate the parameters. Its also notable that
the final mean of the parameters may not be the best point estimator, and that perhaps
using the mode would be a better way to estimate the output. Regardless, the time series'
generated from the estimated mean seems to match the correct signal very well 
(\autoref{fig:FitComparisonLowNoise}).

%HIGH NOISE SECTION, with a signal
\subsection{High Noise}
\label{sec:HighNoise}
\begin{figure}
\includegraphics[clip=true,trim=6cm 2cm 6cm 3.5cm,width=17cm]{images/realization_highnoise}
\caption{Test Signals with high noise compared to the clean signal, $\sigma_x = .01, \sigma_y=.005$}
\label{fig:HighNoiseRealization}
\end{figure}
\begin{figure}
\includegraphics[clip=true,trim=6cm 2cm 6cm 3.5cm,width=17cm]{images/preprocessed_highnoise}
\caption{A comparison of the preprocessed signals for the high noise case.}
\label{fig:PreprocessedHighNoise}
\end{figure}
\begin{figure}
\includegraphics[clip=true,trim=6cm 2cm 6cm 3.5cm,width=17cm]{images/comparison_highnoise}
\caption{A comparison of the fitted signals for the high noise case.}
\label{fig:FitComparisonHighNoise}
\end{figure}
\begin{figure}
\includegraphics[clip=true,trim=6cm 2cm 6cm 3.5cm,width=17cm]{images/highnoise_56_noise}
\caption{Two particular preprocessed noise realizations for the high noise case.}
\label{fig:NoiseComparisonJustTwo}
\end{figure}

@@

For the high noise simulation, the exact same procedure was performed as with the lower
noise simulation except that $\sigma_y$ and $\sigma_x$ were set to $.01$ and $.005$,
respectively. This an order of magnitude higher than the previous tests, and indeed the
noise appears to dominate the output, as seen in \autoref{fig:HighNoiseRealization}.  
As before, the results of preprocessing are helpful in identifying the reason for
bias in the estimated output signal; thus the preprocessed signals are shown in 
\autoref{fig:PreprocessedHighNoise}. The results of the particle filter
for each of the eleven runs are shown in \autoref{fig:FitComparisonHighNoise}. 
As before, the preprocessing led the algorithm
to somewhat higher activation levels, and it would appear that the subtleties of
different time constants are being lost due to the noise. This is easily seen
by observing the post stimulus undershoots; many of the estimated time series
have no post-stimulus undershoot but rather decay directly to the base level.

@@

The parameters arrived at for all ten filter runs are shown in \autoref{tab:HighNoiseResults}.
Clearly the additional noise have resulted in much more sporadic results. This is often the 
result when the particle filter converges too fast, a result of the weighting functions' variance
being significantly smaller than the measurement noise ($.005$ vs. $.01$). The error
certainly suffers due to this effect.

%NO SIGNAL, LOW NOISE
\subsection{Pure-Noise, low magnitude}
\label{sec:PureNoiseLowMag}
The two final single-voxel tests force the particle filter to attempt to learn a noise-only
time series. In the first test the noise used will be the same as that from the \autoref{sec:HighNoise},
$\sigma_x = .01, \sigma_y = .005$. The parameters will be set the same as well, but the
stimulus/input will be set to 0 the entire time. In effect it is a region of the brain for
which no visual stimulus induces activity. For the particle filter, on the other hand,
the stimulus will be left as in the previous two sections. The question then is how the
particle filter will respond to signal that does not correlate with the input, and how
the output may be differentiated from a time series that does. The time series'
are shown in \autoref{fig:NoiseOnly}, the preprocessed versions are shown in \autoref{fig:PreprocessedNoiseOnly}.
The line fits are shown in \autoref{fig:fits_noiseonly}.

\begin{figure}[H]
\centering
\includegraphics[clip=true,trim=6cm 3cm 6cm 3cm,height=9cm]{images/realization_noiseonly}
\caption{Time-series lacking any real signal. With, $\sigma_x = .01, \sigma_y=.005$}
\label{fig:NoiseOnly}
\end{figure}

\begin{figure}[H]
\centering
\includegraphics[clip=true,trim=6cm 3cm 6cm 3cm,height=9cm]{images/preprocessed_noiseonly}
\caption{A comparison of the preprocessed signals for the signal-free case. (
$\sigma_y = .01, \sigma_x = .005$)}
\label{fig:PreprocessedNoiseOnly}
\end{figure}

\begin{table}[t]
\centering
\begin{tabular}{|c | c | c | c | c | c | c | c | c |}
\hline 
$\tau_0$ & $\alpha$ & $E_0$    & $V_0$    & $\tau_s$ & $\tau_f$ & $\epsilon$  &  $\sqrt{MSR}$   \\
\hline 
1.0324 & 0.33211 & 0.34058 & 0.03012 & 1.40665 & 2.52079 & 0.5311 & \\
 0.98189 & 0.33047 & 0.3386 & 0.03014 & 1.45707 & 2.47232 & 0.45049 & 0.00158551 \\
 1.0429 & 0.33224 & 0.34124 & 0.02946 & 1.4618 & 2.49245 & 0.43012 &  0.00165055 \\
 1.02054 & 0.3321 & 0.33484 & 0.02586 & 1.45848 & 2.48741 & 0.4193 &  0.00150604 \\
 1.0565 & 0.33405 & 0.33758 & 0.02791 & 1.43784 & 2.52545 & 0.47517 & 0.00151513 \\
 1.01867 & 0.33528 & 0.33918 & 0.02782 & 1.48345 & 2.49605 & 0.44209 &0.00155645 \\
 1.051 & 0.33038 & 0.33837 & 0.02985 & 1.47651 & 2.48621 & 0.42719 &  0.00158526 \\
 1.00281 & 0.32929 & 0.33988 & 0.0298 & 1.43519 & 2.49256 & 0.48899 & 0.00164299 \\
 1.00893 & 0.33273 & 0.33982 & 0.0289 & 1.42903 & 2.49754 & 0.45688 & 0.0016793  \\
 1.01289 & 0.33275 & 0.3376 & 0.02997 & 1.41188 & 2.49881 & 0.50628 & 0.00182622 \\
 1.10247 & 0.33371 & 0.3419 & 0.02939 & 1.43774 & 2.53384 & 0.44079 & 0.00195009 \\
\hline                                                                
1.03009 & 0.33228 & 0.33905 & 0.02902 & 1.44506 & 2.50031 & 0.46076 & 0.00165129 \\
\hline 
\end{tabular}
\caption{Estimated Parameters on 11 different runs with low noise and no signal present.}
\label{tab:NoiseOnlyResults} 
\end{table}

@@

From the results it is clear that the parameters have not taken any visible turns into
unrealistic territory. It appears individual parameters will not be a distinguishing factor
in determining the validity of the model. Additionally the raw square-root mean-squared-residual
also appears to be lower than the SMSR of the low noise simulation from earlier. 
The precision with which the convergence occurred is actually the best seen in any
simulation.

However, the reason for low error may in fact be simply the fact that the over all
signal is significantly lower than any other simulation seen in other simulations.
The peaks never even cross 1\% difference, with most staying even below .5\% 
(\autoref{fig:PreprocessedNoiseOnly}). Because the signal is generally staying
well within the range of $.005$, which is the standard deviation of the weighting function
its unlikely that any particles will have a substantially different wieght. Thus, the reason
the fits are in such agreement is that no particles gained increased weight, and so
no convergence occurred at all, as further shown by \autoref{fig:JustNoiseConvergence}. 
Thus we have now seen how the particle filter responds when the signal is completely
random and low level, however there is also some possibility that the signal will
be random, yet active. Thus one more test is necessary on
a signal with enough noise to actually be considered a signal.

@@

As before, the convergence all seems to follow a similar path, leaving almost no
variance in the estimated time series (\autoref{fig:fits_noiseonly_high}). Now, however
the peaks are comparable to peaks in a real timeseries. Interestingly the algorithm
suffered from almost constant particle deprevation, meaning that the heuristic
for rescuing the particle filter from particle deprivation, discuessed in
\autoref{sec:Resampling} in fact was working against the proper course of action. Of course
 the proper course of action in this case would be for the particle filter to fail,
 since no particle can really properly estimate a random sequence. When this mechanism
 was removed, all 11 runs stopped due to particle deprevation (all weights hit zero). 
 The problem with allowing particle deprivation to occur is that it can \emph{rarely}
 occur if the prior happens to not be dense enough in the area of the solution. Thus,
 there is an issue of false positives vs. false negatives. A middle ground, such as 
 allowing deprivation to occur if the recovery mechanism has already triggered once
 may be an acceptable compromise, however it would be difficult to strike the 
 right balance. Instead for the purposes here we depend on the residual to 
 evaluate if the results were reasonable or not.

@@

\begin{figure}[H]
\centering
\subfigure[$\tau_0$, $\alpha$, $E_0$, $V_0$]
{\includegraphics[clip=true,trim=6cm 3cm 6cm 3cm, height=9cm]{images/justbignoise_1}}\\
\end{figure}
\begin{figure}[H]
\subfigure[$\tau_s$, $\tau_f$, $\epsilon$, $V$] 
{\includegraphics[clip=true,trim=6cm 3cm 6cm 3cm, height=9cm]{images/justbignoise_2}}\\
\end{figure}
\begin{figure}[H]
\subfigure[$Q$, $S$, $F$, $BOLD$ ]
{\includegraphics[clip=true,trim=6cm 3cm 6cm 3cm, height=9cm]{images/justbignoise_3}}
\caption{Converging histogram for parameters when the signal consists purely of noise, with peaks comparable  to 
the convential BOLD signal. ($\sigma_y = .01, \sigma_x = .005$). Same run as fit in \autoref{fig:justbignoise_fit_0}}
\label{fig:JustNoiseConvergenceLarge}
\end{figure}

From \autoref{fig:justbignoise_fit_0} it is clear that, despite the fact that the input 
is competely detached from the signal, a non-zero response by the particle filter does
lead to decreased MSR. This is primarily due to the preprocessing which forces the signal to
have a non-zero mean. While this scheme does create a type of artificial activation for
the particle filter, in reality the MSR will still be less than if the signal was left
with zero mean, and the particle filter stayed flat. At best the particle filter will 
over-smooth creating a line that does down the center of the noisy signal. In such a case
the MSR will be the same as the alternative.

>>>>

\chapter{Simulation}
\label{sec:SimulationResults}
Two levels of simulation are discussed in this section. First the simulation of single time series
and second, a slice of simulated FMRI time series. The single time series tests
are necessary to investigate the power of particle filters in 
identifying the BOLD model in a noisy environment. If the particle filter is unable
to get back to the \emph{true} time-series in spite of noise then the particle filter
will not be useful. Additionally, determination of reasonable parameters from the
time-series is also desirable. The reason to then simulate an entire slice is that
Physics-Oriented Simulated Scanner for Understanding MRI (POSSUM)
models noise more realistically, for instance by adding noise in the frequency domain. 
It also provides insight into the applicability of particle filters in modeling the BOLD 
signal on a large scale. Both sorts of tests are usefull in determining the
consistency of the particle filter. For the single voxel tests, each test will
actually be performed eleven different times, each with a new noise realization.
If runs fail to have consistent estimates of the underlying BOLD signal, then
this would indicate a large variance error in the model. Similarly, in the
Possum simulation there are actually only four discrete parameter sets driving the time-series.
Nevertheless each voxel will have a different tissue composition and a different
noise realization. Therefore the Possum simulation should also provide a good
test of the consistency of the BOLD model.

\section{Single Time Series}
\label{sec:Single Voxel Simulation}
Given the state-space equations for the BOLD signal, simulating a single time 
series is relatively straight forward. After generating a true signal,
identically and independently distributed (I.I.D.) Gaussian noise and a Wiener
process with Gaussian I.I.D. steps wered added to this true signal. Finally a 
carrier level is added, since BOLD is typically
measured as a \% difference from the baseline. The particle filter
algorithm will immediately remove this by calculating the \% difference, 
but adding a carrier level meant that the exact same algorithm used 
for simulated data could be used for the real data.

Once this noisy simulated time series was generated, the particle filter algorithm
 was then run on single time series image. Here I will include two sets of tests 
to determine the power of the particle filter in modeling. The first set are 
used to demonstrate the ability of the particle filter to find the most likely
set of parameters/state variables over the course of the run. The second type
of test will investigate the problem of false-positives. As the first
round of tests will show, given that there is an underlying BOLD signal, 
the particle filter is excellent at finding the most probable cause of 
the observed signal. However, as discussed in the false positive tests,
even if there is no underlying signal, the particle filter will still determine
the most likely BOLD signal; which may not be flat. Thus, it is 
important to investigate methods of determining false positives using 
other methods.

%LOW NOISE SECTION, with a signal
\subsection{Signal with Low Noise}
\label{sec:SimLowNoise}
%tests had noise of $ \{\sigma_y = .01, \sigma_x = .005\} $, where $\sigma_y$ is the
%measurement noise, and $\sigma_x$ is the wiener step size. Both signals used the
%parameters .
%The particle filter used the parameters defined in \autoref{tab:Prior} (\autoref{sec:PriorDistrib}),
%thus the particle filter was not centered over the correct values. 
\begin{figure}
\includegraphics[clip=true,trim=6cm 2cm 6cm 3.5cm,width=7cm]{images/realization_lownoise}
\caption{Test Signals with low noise compared to the clean signal.}
\label{fig:LowNoiseRealization}
\end{figure}
\begin{figure}
\includegraphics[clip=true,trim=6cm 2cm 6cm 3.5cm,width=7cm]{images/preprocessed_lownoise}
\caption{A comparison of the preprocessed signals for the low noise case. This is the
noisy input to the actual particle filter algorithm.}
\label{fig:PreprocessedLowNoise}
\end{figure}
\begin{figure}
\includegraphics[clip=true,trim=6cm 2cm 6cm 3.5cm,width=17cm]{images/comparison_lownoise}
\caption{A comparison of the fitted signals for the low noise case.}
\label{fig:FitComparisonLowNoise}
\end{figure}

To begin the single voxel simulation; I generated a signal using the following parameters:
$\{\tau_0 = 1.45, \alpha = .3, E_0 = .47, V_0 = .044, \tau_s = 1.94, \tau_f = 1.99, \epsilon = 1.8\}$.
These same parameters were also used in the high noise case. The noise realizations were
then generated based on a measurement noise ($\sigma_y$) of $.001$ and a drift standard deviation
($\sigma_x$) of $.0005$. The measurement noise as well as the steps of the drift
were taken to be Gaussian
For this low noise case, the eleven realizations are shown in \autoref{fig:LowNoiseRealization}.
The actual signal delivered into the particle filter
was the result of some preprocessing to remove drift, as described in 
\autoref{sec:Methods Preprocessing}. The bias introduced into the signal by preprocessing 
may have some effect on the resulting fit; thus the preprocessed signal is compared
to the true BOLD signal in \autoref{fig:PreprocessedLowNoise}.
The preprocessing consists of several steps, which are discussed in detail in \autoref{sec:Methods Preprocessing}.
The data shows that the spline struggles to match the signal near the end, although 
overall, the preprocessing seems to have been successful at removing trends. Finally, the 
set of fits to the true BOLD signal are shown in \autoref{fig:FitComparisonLowNoise}.
The filtering property of the particle filter is clear from the results here.  In several locations the output
of the particle filter looks like a filtered version of the input. For instance toward the
end, the estimates stay flat in spite of the preprocessed data drifting off. By
this point, the algorithm has converged sufficiently to know that that such a movement in
signal isn't possible. A similar circumstance occurs at 100 seconds in. 
It is also worth noting the bias introduced by pre-processing 
results in a slightly larger peak signal. The final parameter sets are shown in 
\autoref{tab:LowNoiseResults}. Note that throughout the results, unless otherwise specified,
the residual will be defined as the square root of the mean squared residual, 
\begin{equation}
\text{Residual} = \sqrt{\sum (\hat{y}_k - y_k)^2}
\end{equation}
where $\hat{y}_k$ is the estimated output at time $k$ and $y_k$ is the output
sample at time $k$. Note the subtle distinction between this and the error
which is defines as 
\begin{equation}
\text{Error} = \sqrt{\sum (\hat{y}_k - Y_k)^2}
\end{equation}
where $\hat{y}_k$ is the estimated output, and $Y_k$ is the underlying (free of noise)
output. Throughout this section and the next, the terms \emph{error} and \emph{residual}
will be referencing these values, for an entire timeseries.

@@

There are a few results worth noting in the final parameter estimates (which 
are the mean of the particle filter's posterior distribution). First the time constants vary 
greatly across
runs with different noise realizations, yet the sum of the individual time constants
($\tau_f$, $\tau_s$ and $\tau_0$) seems to be more consistent. In general the 
time constants fall short of the true time constant. This could be a limitation
based on the prior distribution (which notably has an initial mean below the true values) 
or it could be caused by the disassociated benefits of a correct time
constant with regards to a correct output. It often takes several measurement periods before 
a difference
in time constants becomes relevant to the weight. Its also possible that output is insensitive
to differences in the time constants; althouhg varying the activation durations could induce
better estimates.  The large variation of $V_0$ are also interesting, although as one final
covariance matrix shows, there is a definite negative correlation between $\epsilon$ and $V_0$.
In general,
with this admittedly small amount of noise, it would appear that the relation between a set
of parameters/stimuli and the output is not injective; a time
series is not unique to a single set of parameters. This is good justification that 
simultaneous blood volume or tagged flow calculations with the conventional FMRI 
could benefit the model.

The covariance matrix (\autoref{tab:CovSim}) provides a great deal of insight into 
the way parameters evolve in the particle filter. The fact that there are so many 
covariance elements that are as big or bigger than variance elements indicates just
loose the parameters set is. Every single parameters, including $\alpha$ seems to
have a non-zero covariance. Why $\alpha$ and $\epsilon$ would be positively related
is one of the most interesting element. Given the difficulty in mapping out a 
seven dimensional space, it is difficult to know why such correlations exist. 
 Notice the covariance of $\tau_f$ and $\tau_0$
is $-0.019$ whereas
the variance of $\tau_0$ and $\tau_f$ are $0.019$ and $0.04$ respectively. Clearly there is 
some measure of ill-defined behavior between the different time-constants, which is to
be expected given the number steps before a stimuli affects the output. 
The convergence properties of the first run in \autoref{tab:LowNoiseResults} 
demonstrates the migration of parameters to match the estimates. Given the relatively high
variance of parameters, its likely that more measurements or perhaps more variation
in the stimuli could further differentiate the parameters. Its also notable that
the final mean of the parameters may not be the best point estimator, and that perhaps
using the mode would be a better way to estimate the output. Regardless, the time series'
generated from the estimated mean seems to match the correct signal very well 
(\autoref{fig:FitComparisonLowNoise}).

%HIGH NOISE SECTION, with a signal
\subsection{High Noise}
\label{sec:HighNoise}
\begin{figure}
\includegraphics[clip=true,trim=6cm 2cm 6cm 3.5cm,width=7cm]{images/realization_highnoise}
\caption{Test Signals with high noise compared to the clean signal, $\sigma_x = .01, \sigma_y=.005$}
\label{fig:HighNoiseRealization}
\end{figure}
\begin{figure}
\includegraphics[clip=true,trim=6cm 2cm 6cm 3.5cm,width=7cm]{images/preprocessed_highnoise}
\caption{A comparison of the preprocessed signals for the high noise case.}
\label{fig:PreprocessedHighNoise}
\end{figure}
\begin{figure}
\includegraphics[clip=true,trim=6cm 2cm 6cm 3.5cm,width=17cm]{images/comparison_highnoise}
\caption{A comparison of the fitted signals for the high noise case.}
\label{fig:FitComparisonHighNoise}
\end{figure}
\begin{figure}
\includegraphics[clip=true,trim=6cm 2cm 6cm 3.5cm,width=17cm]{images/highnoise_56_noise}
\caption{Two particular preprocessed noise realizations for the high noise case.}
\label{fig:NoiseComparisonJustTwo}
\end{figure}

@@

For the high noise simulation, the exact same procedure was performed as with the lower
noise simulation except that $\sigma_y$ and $\sigma_x$ were set to $.01$ and $.005$,
respectively. This is an order of magnitude higher than the previous tests, and indeed the
noise appears to dominate the output, as seen in \autoref{fig:HighNoiseRealization}.  
As before, the results of preprocessing are helpful in identifying the reason for
bias in the estimated output signal; thus the preprocessed signals are shown in 
\autoref{fig:PreprocessedHighNoise}. The results of the particle filter
for each of the eleven runs are shown in \autoref{fig:FitComparisonHighNoise}. 
As before, the preprocessing led the algorithm
to somewhat higher activation levels, and it would appear that the subtleties of
different time constants are being lost due to the noise. This is easily seen
by observing the post stimulus undershoots; many of the estimated time series
have no post-stimulus undershoot but rather decay directly to the base level.

@@

The parameters arrived at for all ten filter runs are shown in \autoref{tab:HighNoiseResults}.
Clearly the additional noise have resulted in much more sporadic results. This is often the 
result when the particle filter converges too fast, a result of the weighting functions' variance
being smaller than the measurement noise ($.005$ vs. $.01$). The error
certainly suffers due to this effect.

%NO SIGNAL, LOW NOISE
\subsection{Pure-Noise, low magnitude}
\label{sec:PureNoiseLowMag}
The two final single-voxel tests force the particle filter to attempt to learn a noise-only
time series. In the first test the noise used will be the same as that from the \autoref{sec:HighNoise},
$\sigma_x = .01, \sigma_y = .005$. The parameters will be set the same as well, but the
stimulus/input neuronal efficiency ($\epsilon$) will be set to 0. In effect it is a region of the brain for
which no visual stimulus induces activity. The question then is how the
particle filter will respond to signal that does not correlate with the input, and how
the output may be differentiated from a time series that does. The time series'
are shown in \autoref{fig:NoiseOnly}, the preprocessed versions are shown in \autoref{fig:PreprocessedNoiseOnly}.
The line fits are shown in \autoref{fig:fits_noiseonly}.

\begin{figure}[H]
\centering
\includegraphics[clip=true,trim=6cm 3cm 6cm 3cm,height=7cm]{images/realization_noiseonly}
\caption{Time-series lacking any real signal. With, $\sigma_x = .01, \sigma_y=.005$}
\label{fig:NoiseOnly}
\end{figure}

\begin{figure}[H]
\centering
\includegraphics[clip=true,trim=6cm 3cm 6cm 3cm,height=7cm]{images/preprocessed_noiseonly}
\caption{A comparison of the preprocessed signals for the signal-free case. (
$\sigma_y = .01, \sigma_x = .005$)}
\label{fig:PreprocessedNoiseOnly}
\end{figure}

\begin{table}[t]
\centering
\begin{tabular}{|c | c | c | c | c | c | c | c | c |}
\hline 
$\tau_0$ & $\alpha$ & $E_0$    & $V_0$    & $\tau_s$ & $\tau_f$ & $\epsilon$  &  $\sqrt{MSR}$   \\
\hline 
1.0324 & 0.33211 & 0.34058 & 0.03012 & 1.40665 & 2.52079 & 0.5311 & todo \\
 0.98189 & 0.33047 & 0.3386 & 0.03014 & 1.45707 & 2.47232 & 0.45049 & 0.00158551 \\
 1.0429 & 0.33224 & 0.34124 & 0.02946 & 1.4618 & 2.49245 & 0.43012 &  0.00165055 \\
 1.02054 & 0.3321 & 0.33484 & 0.02586 & 1.45848 & 2.48741 & 0.4193 &  0.00150604 \\
 1.0565 & 0.33405 & 0.33758 & 0.02791 & 1.43784 & 2.52545 & 0.47517 & 0.00151513 \\
 1.01867 & 0.33528 & 0.33918 & 0.02782 & 1.48345 & 2.49605 & 0.44209 &0.00155645 \\
 1.051 & 0.33038 & 0.33837 & 0.02985 & 1.47651 & 2.48621 & 0.42719 &  0.00158526 \\
 1.00281 & 0.32929 & 0.33988 & 0.0298 & 1.43519 & 2.49256 & 0.48899 & 0.00164299 \\
 1.00893 & 0.33273 & 0.33982 & 0.0289 & 1.42903 & 2.49754 & 0.45688 & 0.0016793  \\
 1.01289 & 0.33275 & 0.3376 & 0.02997 & 1.41188 & 2.49881 & 0.50628 & 0.00182622 \\
 1.10247 & 0.33371 & 0.3419 & 0.02939 & 1.43774 & 2.53384 & 0.44079 & 0.00195009 \\
\hline                                                                
1.03009 & 0.33228 & 0.33905 & 0.02902 & 1.44506 & 2.50031 & 0.46076 & 0.00165129 \\
\hline 
\end{tabular}
\caption{Estimated Parameters on 11 different runs with low noise and no signal present.}
\label{tab:NoiseOnlyResults} 
\end{table}

@@

The data shows that the parameters have not taken any visible turns into
unrealistic territory; individual parameters will not be a distinguishing factor
in determining the validity of the model. Additionally the residual 
also appears to be lower than the residual of the low noise simulation from earlier. 
The precision with which the convergence occurred is actually the best seen in any
simulation.  The reason for the low error is that the overall
signal is significantly smaller than any other simulation in this thesis.
The peaks never even 1\% difference, with most staying even below .5\% 
(\autoref{fig:PreprocessedNoiseOnly}). Because the stayed 
well within the range of $.005$, which is the standard deviation of the weighting function
the particles did not ever get substantially different weights. Thus, the reason
the fits are in such agreement is that no particles gained increased weight, and so
no convergence occurred at all, as further shown by \autoref{fig:JustNoiseConvergence}. 
This is good evidence that a per time-series weighting function could benefit
even out the convergence rate; and thus allow better determination between signals
like this and low noise case from earlier.
Thus we have now seen how the particle filter responds when the signal is completely
random and low level, however there is also some possibility that the signal will
be random, yet active. Thus one more test is necessary; time series with no 
underlying signal (as far as responding stimuli) but with enough noise to 
actually be considered a signal.

@@

As before, the convergence all seems to follow a similar path, leaving almost no
variance in the estimated time series (\autoref{fig:fits_noiseonly_high}). Now, however
the peaks are comparable to peaks in a real timeseries. Interestingly the algorithm
suffered from almost constant particle deprivation, meaning that the heuristic
for rescuing the particle filter from particle deprivation, discussed in
\autoref{sec:Resampling} in fact was working against the proper course of action. 
 The proper course of action in this case would be for the particle filter to fail,
 since no particle can really properly estimate a random sequence. When this mechanism
 was removed, all 11 runs stopped due to particle deprivation (all weights hit zero). 
 The problem with allowing particle deprivation to occur is that it can \emph{rarely}
 occur if the prior happens to not be dense enough in the area of the solution. Thus,
 there is an issue of false positives vs. false negatives. A middle ground, such as 
 allowing deprivation to occur if the recovery mechanism has already triggered once
 may be an acceptable compromise, however it would be difficult to strike the 
 right balance. Instead, for the purposes here, I depend on other methods, described
 later, evaluate if the results were reasonable or not.

@@

%\begin{figure}[H]
%\centering
%\subfigure[$\tau_0$, $\alpha$, $E_0$, $V_0$]
%{\includegraphics[clip=true,trim=6cm 3cm 6cm 3cm, height=9cm]{images/justbignoise_1}}\\
%\end{figure}
%\begin{figure}[H]
%\subfigure[$\tau_s$, $\tau_f$, $\epsilon$, $V$] 
%{\includegraphics[clip=true,trim=6cm 3cm 6cm 3cm, height=9cm]{images/justbignoise_2}}\\
%\end{figure}
%\begin{figure}[H]
%\subfigure[$Q$, $S$, $F$, $BOLD$ ]
%{\includegraphics[clip=true,trim=6cm 3cm 6cm 3cm, height=9cm]{images/justbignoise_3}}
%\caption{Converging histogram for parameters when the signal consists purely of noise, with peaks comparable  to 
%the convential BOLD signal. ($\sigma_y = .01, \sigma_x = .005$). Same run as fit in \autoref{fig:justbignoise_fit_0}}
%\label{fig:JustNoiseConvergenceLarge}
%\end{figure}

From \autoref{fig:justbignoise_fit_0} it is clear that, despite the fact that the input 
is competely detached from the signal, a non-zero response by the particle filter does
lead to decreased residual. This is primarily due to the preprocessing which forces the signal to
have a non-zero mean. While this scheme could create a type of artificial activation for
the particle filter, in reality the residual will still be less than if the signal were left
with zero mean, and the particle filter stayed flat. At best the particle filter will 
over-smooth, estimating a centerline to noisy signal. In such a case
the residual will be the same as the alternative, although the mutual information will 
be different.
